---
title: "Climate Change Projections: CNRM-ESM2-1"
author: "Elmera Azadpour"
date: "12/21/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(lubridate)
library(ggpubr)
library(zoo)
```

## read in climate data exported from jupyter notebook, code can be found here: https://github.com/aloha-aina/Aloha-Aina-Master 
```{r}
historical = read.csv(here("climate_change_data","oahu_historical_2015_2100.csv")) %>%  rename(date = dayofyear, pcp = pr) 
ssp126 = read.csv(here("climate_change_data","oahu_ssp126_2015_2100.csv")) %>% 
  rename(date = dayofyear, pcp = pr) 
ssp245 = read.csv(here("climate_change_data","oahu_ssp245_2015_2100.csv")) %>% 
  rename(date = dayofyear, pcp = pr) 
ssp370 = read.csv(here("climate_change_data","oahu_ssp370_2015_2100.csv")) %>% 
  rename(date = dayofyear, pcp = pr)
ssp585 = read.csv(here("climate_change_data","oahu_ssp585_2015_2100.csv")) %>% 
  rename(date = dayofyear, pcp = pr)

## convert date column to year, month, day components using luberdate package
historical$date <- ymd(historical$date)
ssp126$date <- ymd(ssp126$date)
ssp245$date <- ymd(ssp245$date)
ssp370$date <- ymd(ssp370$date)
ssp585$date <- ymd(ssp585$date)
```

## explore daily avg pcp data from jupyter notebook 
```{r}
daily_avg_proj <- 
  ggplot() +
  geom_line(data = ssp126, aes(x=date, y=pcp, color="orange"), size = 0.5) + 
  geom_line(data = ssp245, aes(x=date, y=pcp, color="red"), size = 0.5) +
  geom_line(data = ssp370, aes(x=date, y=pcp, color = "blue"), size = 0.5) +
  geom_line(data = ssp585, aes(x=date, y=pcp, color = "seagreen"), size = 0.5) + 
  theme_minimal() + 
  scale_y_continuous(breaks=seq(0,7,1)) +
  scale_x_date(date_breaks = "10 years", date_labels = "%Y") +
  ggtitle("Oahu, HI Daily Avg Precipitation, CNRM-ESM2-1: 2015-2100") +
   labs(x = "Year",
       y = "Precipitation (in)") +
  scale_color_manual(values = c("orange", "red", "blue", "seagreen"), labels = c("spp126", "ssp245", "ssp370", "ssp585"))

daily_avg_proj

## extra bit for regression 
# +
#   geom_smooth(method='lm', formula = y~x) +
#   stat_regline_equation(label.x = 2015,
#                         label.y = 6) +
#   stat_cor(label.x = 10,
#            label.y = 4)
```

## wrangling bits to group_by()
```{r}
# add seperate columns for year, month and day
historical <- historical %>% 
  mutate(year = year(date)) %>% 
  mutate(month = month(date)) %>% 
  mutate(day = day(date))

ssp126 <- ssp126 %>% 
  mutate(year = year(date)) %>% 
  mutate(month = month(date)) %>% 
  mutate(day = day(date))

ssp245 <- ssp245 %>% 
  mutate(year = year(date)) %>% 
  mutate(month = month(date)) %>% 
  mutate(day = day(date))

ssp370 <- ssp370 %>% 
  mutate(year = year(date)) %>% 
  mutate(month = month(date)) %>% 
  mutate(day = day(date))

ssp585 <- ssp585  %>% 
  mutate(year = year(date)) %>% 
  mutate(month = month(date)) %>% 
  mutate(day = day(date))


# group by yearly sum of pcp 
historical_yearly <- historical %>% 
  group_by(year) %>% 
  mutate(sum = sum(pcp)) %>% 
  select(year, sum) %>% 
  distinct()

historical_yearly$year <- as.numeric(historical_yearly$year) 


# group by yearly sum of pcp 
ssp126_yearly <- ssp126 %>% 
  group_by(year) %>% 
  mutate(sum = sum(pcp)) %>% 
  select(year, sum) %>% 
  distinct()

ssp126_yearly$year <- as.numeric(ssp126_yearly$year) 

# group by yearly sum of pcp 
ssp245_yearly <- ssp245 %>% 
  group_by(year) %>% 
  mutate(sum = sum(pcp)) %>% 
  select(year, sum) %>% 
  distinct()

ssp245_yearly$year <- as.numeric(ssp245_yearly$year) 

# group by yearly sum of pcp 
ssp370_yearly <- ssp370 %>% 
  group_by(year) %>% 
  mutate(sum = sum(pcp)) %>% 
  select(year, sum) %>% 
  distinct()

ssp370_yearly$year <- as.numeric(ssp370_yearly$year) 

# group by yearly sum of pcp 
ssp585_yearly <- ssp585 %>% 
  group_by(year) %>% 
  mutate(sum = sum(pcp)) %>% 
  select(year, sum) %>% 
  distinct()

ssp585_yearly$year <- as.numeric(ssp585_yearly$year) 


# group by month (and associated year) sum of pcp
historical_monthly <- historical %>% 
  group_by(month,year) %>% 
  mutate(sum = sum(pcp)) %>% 
  select(month, year, sum) %>% 
  distinct()

historical_monthly$date <- format(as.Date(paste(historical_monthly$year, historical_monthly$month, 01), "%Y %m %d"), "%m-%Y")

historical_monthly$date <- my(historical_monthly$date)


# group by month (and associated year) sum of pcp
ssp126_monthly <- ssp126 %>% 
  group_by(month,year) %>% 
  mutate(sum = sum(pcp)) %>% 
  select(month, year, sum) %>% 
  distinct()

ssp126_monthly$date <- format(as.Date(paste(ssp126_monthly$year, ssp126_monthly$month, 01), "%Y %m %d"), "%m-%Y")

ssp126_monthly$date <- my(ssp126_monthly$date)


# group by month (and associated year) sum of pcp
ssp245_monthly <- ssp245 %>% 
  group_by(month,year) %>% 
  mutate(sum = sum(pcp)) %>% 
  select(month, year, sum) %>% 
  distinct()

ssp245_monthly$date <- format(as.Date(paste(ssp245_monthly$year, ssp245_monthly$month, 01), "%Y %m %d"), "%m-%Y")

ssp245_monthly$date <- my(ssp245_monthly$date)

# group by yearly sum of daily avg pcp 
ssp370_monthly <- ssp370 %>% 
  group_by(month,year) %>% 
  mutate(sum = sum(pcp)) %>% 
  select(month, year, sum) %>% 
  distinct() 
  
ssp370_monthly$date <- format(as.Date(paste(ssp370_monthly$year, ssp370_monthly$month, 01), "%Y %m %d"), "%m-%Y")

ssp370_monthly$date <- my(ssp370_monthly$date)

# group by yearly sum of daily avg pcp 
ssp585_monthly <- ssp585 %>% 
  group_by(month,year) %>% 
  mutate(sum = sum(pcp)) %>% 
  select(month, year, sum) %>% 
  distinct() 
  
ssp585_monthly$date <- format(as.Date(paste(ssp585_monthly$year, ssp585_monthly$month, 01), "%Y %m %d"), "%m-%Y")

ssp585_monthly$date <- my(ssp585_monthly$date)

```


## yearly projected precipiataion analysis 
```{r}
## all three ssps 
yearly_avg_proj <- 
  ggplot() +
  geom_line(data = ssp126_yearly, aes(x=year, y=sum, color="orange"), size = 0.5) + 
  geom_line(data = ssp245_yearly, aes(x=year, y=sum, color="red"), size = 0.5) +
  geom_line(data = ssp370_yearly, aes(x=year, y=sum, color = "blue"), size = 0.5) +
  geom_line(data = ssp585_yearly, aes(x=year, y=sum, color = "seagreen"), size = 0.5) +
  theme_minimal() + 
  scale_y_continuous(breaks=seq(0,50,10)) +
  scale_x_continuous(breaks=seq(2015,2100,10)) + 
  # scale_x_date(date_breaks = "5 years", date_labels = "%Y") +
  ggtitle("Oahu, HI Yearly Sum of Daily Avg Precipitation, CNRM-ESM2-1: 2015-2100") + 
  scale_color_manual(values = c("orange", "red", "blue", "seagreen"), labels = c("ssp126", "ssp245", "ssp370", "ssp585")) +
  labs(x = "Year",
       y = "Precipitation (in)") +
  geom_smooth(data = ssp126_yearly, aes(x=year, y=sum), method='lm', formula = y~x, color = "orange") + 
  geom_smooth(data = ssp245_yearly, aes(x=year, y=sum), method='lm', formula = y~x, color = "red") + 
  geom_smooth(data = ssp370_yearly, aes(x=year, y=sum), method='lm', formula = y~x, color = "blue") + 
  geom_smooth(data = ssp585_yearly, aes(x=year, y=sum), method='lm', formula = y~x, color = "seagreen") + 
  stat_regline_equation(label.y = 45) +
  stat_cor(label.y = 40)

yearly_avg_proj

## historical 
historical_yearly_plot <- 
  ggplot(data = historical_yearly, aes(x=year, y=sum)) +
  geom_line(size = 0.5,color="purple") +
  theme_minimal() + 
  scale_y_continuous(breaks=seq(0,50,10)) +
  scale_x_continuous(breaks=seq(1850,2015,10)) + 
  ggtitle("Oahu, HI Yearly Sum of Daily Avg Precipitation, historical, CNRM-ESM2-1: 1850-2014") + 
  geom_smooth(method='lm', formula = y~x, color = "purple") +
  stat_regline_equation(label.y = 45) +
  stat_cor(label.y = 40) +
   labs(x = "Year",
       y = "Precipitation (in)")

historical_yearly_plot


## ssp126
ssp126_yearly_plot <- 
  ggplot(data = ssp126_yearly, aes(x=year, y=sum)) +
  geom_line(size = 0.5,color="orange") +
  theme_minimal() + 
  scale_y_continuous(breaks=seq(0,50,10)) +
  scale_x_continuous(breaks=seq(2015,2100,10)) + 
  ggtitle("Oahu, HI Yearly Sum of Daily Avg Precipitation,ssp126, CNRM-ESM2-1: 2015-2100") + 
  geom_smooth(method='lm', formula = y~x, color = "orange") +
  stat_regline_equation(label.y = 45) +
  stat_cor(label.y = 40) +
   labs(x = "Year",
       y = "Precipitation (in)")

ssp126_yearly_plot


## ssp245
ssp245_yearly_plot <- 
  ggplot(data = ssp245_yearly, aes(x=year, y=sum)) +
  geom_line(size = 0.5,color="red") +
  theme_minimal() + 
  scale_y_continuous(breaks=seq(0,50,10)) +
  scale_x_continuous(breaks=seq(2015,2100,10)) + 
  ggtitle("Oahu, HI Yearly Sum of Daily Avg Precipitation,ssp245, CNRM-ESM2-1: 2015-2100") + 
  geom_smooth(method='lm', formula = y~x, color = "red") +
  stat_regline_equation(label.y = 45) +
  stat_cor(label.y = 40) +
   labs(x = "Year",
       y = "Precipitation (in)")

ssp245_yearly_plot
  
## ssp370
ssp370_yearly_plot <- 
  ggplot(data = ssp370_yearly, aes(x=year, y=sum)) +
  geom_line(size = 0.5,color="blue") +
  theme_minimal() + 
  scale_y_continuous(breaks=seq(0,50,10)) +
  scale_x_continuous(breaks=seq(2015,2100,10)) + 
  ggtitle("Oahu, HI Yearly Sum of Daily Avg Precipitation, ssp370, CNRM-ESM2-1: 2015-2100") + 
  geom_smooth(method='lm', formula = y~x, color = "blue") +
  stat_regline_equation(label.y = 45) +
  stat_cor(label.y = 40) +
   labs(x = "Year",
       y = "Precipitation (in)")

ssp370_yearly_plot

## ssp585
ssp585_yearly_plot <- 
  ggplot(data = ssp585_yearly, aes(x=year, y=sum)) +
  geom_line(size = 0.5,color="seagreen") +
  theme_minimal() + 
  scale_y_continuous(breaks=seq(0,50,10)) +
  scale_x_continuous(breaks=seq(2015,2100,10)) + 
  ggtitle("Oahu, HI Yearly Sum of Daily Avg Precipitation, ssp585, CNRM-ESM2-1: 2015-2100") + 
  geom_smooth(method='lm', formula = y~x, color = "seagreen") +
  stat_regline_equation(label.y = 45) +
  stat_cor(label.y = 40) +
   labs(x = "Year",
       y = "Precipitation (in)")

ssp585_yearly_plot
```

## month by year projected precipiation analysis
```{r}
## historical monthly 
historical_monthly_plot <- 
  ggplot(data = historical_monthly, aes(x=date, y=sum)) +
  geom_line(size = 0.5,color="purple") +
  theme_minimal() + 
  scale_y_continuous(breaks=seq(0,20,5)) +
  #scale_x_continuous(breaks=seq(2015,2100,10)) + 
  ggtitle("Oahu, HI Monthly Sum of Daily Avg Precipitation, ssp126, CNRM-ESM2-1: 2015-2100") + 
  geom_smooth(method='lm', formula = y~x, color = "purple") +
  stat_regline_equation(label.y = 16) +
  stat_cor(label.y = 15) +
   labs(x = "Year",
       y = "Precipitation (in)")

historical_monthly_plot


## ssp126 monthly 
ssp126_monthly_plot <- 
  ggplot(data = ssp126_monthly, aes(x=date, y=sum)) +
  geom_line(size = 0.5,color="orange") +
  theme_minimal() + 
  scale_y_continuous(breaks=seq(0,20,5)) +
  #scale_x_continuous(breaks=seq(2015,2100,10)) + 
  ggtitle("Oahu, HI Monthly Sum of Daily Avg Precipitation, ssp126, CNRM-ESM2-1: 2015-2100") + 
  geom_smooth(method='lm', formula = y~x, color = "orange") +
  stat_regline_equation(label.y = 16) +
  stat_cor(label.y = 15) +
   labs(x = "Year",
       y = "Precipitation (in)")

ssp126_monthly_plot

## ssp245 monthly 
ssp245_monthly_plot <- 
  ggplot(data = ssp245_monthly, aes(x=date, y=sum)) +
  geom_line(size = 0.5,color="red") +
  theme_minimal() + 
  scale_y_continuous(breaks=seq(0,20,5)) +
  #scale_x_continuous(breaks=seq(2015,2100,10)) + 
  ggtitle("Oahu, HI Monthly Sum of Daily Avg Precipitation, ssp245, CNRM-ESM2-1: 2015-2100") + 
  geom_smooth(method='lm', formula = y~x, color = "red") +
  stat_regline_equation(label.y = 16) +
  stat_cor(label.y = 15) +
   labs(x = "Year",
       y = "Precipitation (in)")

ssp245_monthly_plot

## ssp370 monthly
ssp370_monthly_plot <- 
  ggplot(data = ssp370_monthly, aes(x=date, y=sum)) +
  geom_line(size = 0.5,color="blue") +
  theme_minimal() + 
  scale_y_continuous(breaks=seq(0,20,5)) +
  #scale_x_continuous(breaks=seq(2015,2100,10)) + 
  ggtitle("Oahu, HI Monthly Sum of Daily Avg Precipitation, ssp370, CNRM-ESM2-1: 2015-2100") + 
  geom_smooth(method='lm', formula = y~x, color = "blue") +
  stat_regline_equation(label.y = 16) +
  stat_cor(label.y = 15) +
   labs(x = "Year",
       y = "Precipitation (in)")

ssp370_monthly_plot

## ssp585 monthly
ssp585_monthly_plot <- 
  ggplot(data = ssp585_monthly, aes(x=date, y=sum)) +
  geom_line(size = 0.5,color="seagreen") +
  theme_minimal() + 
  scale_y_continuous(breaks=seq(0,20,5)) +
  #scale_x_continuous(breaks=seq(2015,2100,10)) + 
  ggtitle("Oahu, HI Monthly Sum of Daily Avg Precipitation, ssp585, CNRM-ESM2-1: 2015-2100") + 
  geom_smooth(method='lm', formula = y~x, color = "seagreen") +
  stat_regline_equation(label.y = 16) +
  stat_cor(label.y = 15) + 
   labs(x = "Year",
       y = "Precipitation (in)")

ssp585_monthly_plot
```
## monthly combined projected precipation analysis 
```{r}
monthly_avg_proj <- 
  ggplot() +
  geom_line(data = ssp126_monthly, aes(x=date, y=sum, color= "orange"), size = 0.5) + 
  geom_line(data = ssp245_monthly, aes(x=date, y=sum, color= "red"), size = 0.5) +
  geom_line(data = ssp370_monthly, aes(x=date, y=sum, color = "blue"), size = 0.5) +
  geom_line(data = ssp585_monthly, aes(x=date, y=sum, color = "seagreen"), size = 0.5) +
  theme_minimal() + 
  scale_y_continuous(breaks=seq(0,20,5)) +
  #scale_x_continuous(breaks=seq(2015,2100,10)) + 
  # scale_x_date(date_breaks = "5 years", date_labels = "%Y") +
  ggtitle("Oahu, HI Monthly Sum of Daily Avg Precipitation, CNRM-ESM2-1: 2015-2100") + 
  scale_color_manual(values = c("orange", "red", "blue", "seagreen"), labels = c("ssp126", "ssp245", "ssp370", "ssp585")) +
  labs(x = "Year",
       y = "Precipitation (in)") 

monthly_avg_proj
```

## export graphs 
```{r}
# ggsave("img/daily_avg_proj.png", daily_avg_proj, height = 5, width = 7, dpi = 320)
# ggsave("img/yearly_avg_proj.png", yearly_avg_proj, height = 5, width = 7, dpi = 320)
# ggsave("img/monthly_avg_proj.png", monthly_avg_proj, height = 5, width = 7, dpi = 320)
# ggsave("img/ssp245_yearly_plot.png", ssp245_yearly_plot, height = 5, width = 7, dpi = 320)
# ggsave("img/ssp370_yearly_plot.png", ssp370_yearly_plot, height = 5, width = 7, dpi = 320)
# ggsave("img/ssp585_yearly_plot.png", ssp585_yearly_plot, height = 5, width = 7, dpi = 320)
# ggsave("img/ssp245_monthly_plot.png", ssp245_monthly_plot, height = 5, width = 7, dpi = 320)
# ggsave("img/ssp370_monthly_plot.png", ssp370_monthly_plot, height = 5, width = 7, dpi = 320)
# ggsave("img/ssp585_monthly_plot.png", ssp585_monthly_plot, height = 5, width = 7, dpi = 320)
```

## Lets filter analysis for projections until 2050 instead of 2100 to see how much regression is altered.
```{r}
ssp126_2050 <- ssp126 %>% 
  filter(year %in% (2015:2050))

ssp245_2050 <- ssp245 %>% 
  filter(year %in% (2015:2050))

ssp370_2050 <- ssp370 %>% 
  filter(year %in% (2015:2050))

ssp585_2050 <- ssp585 %>% 
  filter(year %in% (2015:2050))



## ssp126  til 2050
ssp126_2050_monthly_plot <- 
  ggplot(data = ssp126_2050, aes(x=date, y=pcp)) +
  geom_line(size = 0.5,color="orange") +
  theme_minimal() + 
  scale_y_continuous(breaks=seq(0,7,1)) +
  #scale_x_continuous(breaks=seq(2015,2100,10)) + 
  ggtitle("Oahu, HI Daily Avg Precipitation, ssp126, CNRM-ESM2-1: 2015-2050") + 
  geom_smooth(method='lm', formula = y~x, color = "orange") +
  stat_regline_equation(label.y = 6) +
  stat_cor(label.y = 5) +
   labs(x = "Year",
       y = "Precipitation (in)")

ssp126_2050_monthly_plot

## ssp245  til 2050
ssp245_2050_monthly_plot <- 
  ggplot(data = ssp245_2050, aes(x=date, y=pcp)) +
  geom_line(size = 0.5,color="red") +
  theme_minimal() + 
  scale_y_continuous(breaks=seq(0,7,1)) +
  #scale_x_continuous(breaks=seq(2015,2100,10)) + 
  ggtitle("Oahu, HI Daily Avg Precipitation, ssp245, CNRM-ESM2-1: 2015-2050") + 
  geom_smooth(method='lm', formula = y~x, color = "red") +
  stat_regline_equation(label.y = 6) +
  stat_cor(label.y = 5) +
   labs(x = "Year",
       y = "Precipitation (in)")

ssp245_2050_monthly_plot

## ssp370  til 2050
ssp370_2050_monthly_plot <- 
  ggplot(data = ssp370_2050, aes(x=date, y=pcp)) +
  geom_line(size = 0.5,color="blue") +
  theme_minimal() + 
  scale_y_continuous(breaks=seq(0,7,1)) +
  #scale_x_continuous(breaks=seq(2015,2100,10)) + 
  ggtitle("Oahu, HI Daily Avg Precipitation, ssp370, CNRM-ESM2-1: 2015-2050") + 
  geom_smooth(method='lm', formula = y~x, color = "blue") +
  stat_regline_equation(label.y = 6) +
  stat_cor(label.y = 5) +
   labs(x = "Year",
       y = "Precipitation (in)")

ssp370_2050_monthly_plot

## ssp585 monthly til 2050
ssp585_2050_monthly_plot <- 
  ggplot(data = ssp585_2050, aes(x=date, y=pcp)) +
  geom_line(size = 0.5,color="seagreen") +
  theme_minimal() + 
  scale_y_continuous(breaks=seq(0,7,1)) +
  #scale_x_continuous(breaks=seq(2015,2100,10)) + 
  ggtitle("Oahu, HI Daily Avg Precipitation, ssp585, CNRM-ESM2-1: 2015-2050") + 
  geom_smooth(method='lm', formula = y~x, color = "seagreen") +
  stat_regline_equation(label.y = 6) +
  stat_cor(label.y = 5) + 
   labs(x = "Year",
       y = "Precipitation (in)")

ssp585_2050_monthly_plot

```


## yealy regression til 2050 
```{r}
# group by yearly sum of pcp 
ssp126_2050_yearly <- ssp126_2050 %>% 
  group_by(year) %>% 
  mutate(sum = sum(pcp)) %>% 
  select(year, sum) %>% 
  distinct()

ssp126_2050_yearly$year <- as.numeric(ssp126_2050_yearly$year) 

# group by yearly sum of pcp 
ssp245_2050_yearly <- ssp245_2050 %>% 
  group_by(year) %>% 
  mutate(sum = sum(pcp)) %>% 
  select(year, sum) %>% 
  distinct()

ssp245_2050_yearly$year <- as.numeric(ssp245_2050_yearly$year) 

# group by yearly sum of pcp 
ssp370_2050_yearly <- ssp370_2050 %>% 
  group_by(year) %>% 
  mutate(sum = sum(pcp)) %>% 
  select(year, sum) %>% 
  distinct()

ssp370_2050_yearly$year <- as.numeric(ssp370_2050_yearly$year) 

# group by yearly sum of pcp 
ssp585_2050_yearly <- ssp585_2050 %>% 
  group_by(year) %>% 
  mutate(sum = sum(pcp)) %>% 
  select(year, sum) %>% 
  distinct()

ssp585_2050_yearly$year <- as.numeric(ssp585_2050_yearly$year) 



## ssp126 2050 plot
ssp126_2050_yearly_plot <- 
  ggplot(data = ssp126_2050_yearly, aes(x=year, y=sum)) +
  geom_line(size = 0.5,color="orange") +
  theme_minimal() + 
  scale_y_continuous(breaks=seq(0,50,10)) +
  scale_x_continuous(breaks=seq(2015,2050,10)) + 
  ggtitle("Oahu, HI Yearly Sum of Daily Avg Precipitation,ssp126, CNRM-ESM2-1: 2015-2050") + 
  geom_smooth(method='lm', formula = y~x, color = "orange") +
  stat_regline_equation(label.y = 45) +
  stat_cor(label.y = 40) +
   labs(x = "Year",
       y = "Precipitation (in)")

ssp126_2050_yearly_plot


## ssp245 2050 plot 
ssp245_2050_yearly_plot <- 
  ggplot(data = ssp245_2050_yearly, aes(x=year, y=sum)) +
  geom_line(size = 0.5,color="red") +
  theme_minimal() + 
  scale_y_continuous(breaks=seq(0,50,10)) +
  scale_x_continuous(breaks=seq(2015,2050,10)) + 
  ggtitle("Oahu, HI Yearly Sum of Daily Avg Precipitation, ssp245, CNRM-ESM2-1: 2015-2050") + 
  geom_smooth(method='lm', formula = y~x, color = "red") +
  stat_regline_equation(label.y = 45) +
  stat_cor(label.y = 40) +
   labs(x = "Year",
       y = "Precipitation (in)")

ssp245_2050_yearly_plot
  
## ssp370 2050 plot 
ssp370_2050_yearly_plot <- 
  ggplot(data = ssp370_2050_yearly, aes(x=year, y=sum)) +
  geom_line(size = 0.5, color="blue") +
  theme_minimal() + 
  scale_y_continuous(breaks=seq(0,50,10)) +
  scale_x_continuous(breaks=seq(2015,2050,10)) + 
  ggtitle("Oahu, HI Yearly Sum of Daily Avg Precipitation, ssp370, CNRM-ESM2-1: 2015-2050") + 
  geom_smooth(method='lm', formula = y~x, color = "blue") +
  stat_regline_equation(label.y = 45) +
  stat_cor(label.y = 40) +
   labs(x = "Year",
       y = "Precipitation (in)")

ssp370_2050_yearly_plot

## ssp585 2050 plot 
ssp585_2050_yearly_plot <- 
  ggplot(data = ssp585_2050_yearly, aes(x=year, y=sum)) +
  geom_line(size = 0.5,color="seagreen") +
  theme_minimal() + 
  scale_y_continuous(breaks=seq(0,50,10)) +
  scale_x_continuous(breaks=seq(2015,2050,10)) + 
  ggtitle("Oahu, HI Yearly Sum of Daily Avg Precipitation, ssp585, CNRM-ESM2-1: 2015-2050") + 
  geom_smooth(method='lm', formula = y~x, color = "seagreen") +
  stat_regline_equation(label.y = 45) +
  stat_cor(label.y = 40) +
   labs(x = "Year",
       y = "Precipitation (in)")

ssp585_2050_yearly_plot

```

