---
title: 'Climate Change Projections: CNRM-ESM2-1'
author: "Elmera Azadpour"
date: "12/21/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(lubridate)
library(ggpubr)
library(zoo)
```

## read in daily avg and total climate data projection and historical data exported from jupyter notebook, code can be found here: https://github.com/aloha-aina/Aloha-Aina-Master 
```{r}
## Note: these data are from climate model CNRM-ESM2-1 (France)
## daily avgs
historical_avg = read.csv(here("climate_change_data", "CNRM_daily_avg","oahu_historical_1850_2014_avg.csv")) %>%  
  rename(date = dayofyear, pcp = pr) 
ssp126_avg = read.csv(here("climate_change_data", "CNRM_daily_avg","oahu_ssp126_2015_2100_avg.csv")) %>% 
  rename(date = dayofyear, pcp = pr) 
ssp245_avg = read.csv(here("climate_change_data","CNRM_daily_avg","oahu_ssp245_2015_2100_avg.csv")) %>% 
  rename(date = dayofyear, pcp = pr) 
ssp370_avg = read.csv(here("climate_change_data","CNRM_daily_avg","oahu_ssp370_2015_2100_avg.csv")) %>% 
  rename(date = dayofyear, pcp = pr)
ssp585_avg = read.csv(here("climate_change_data","CNRM_daily_avg","oahu_ssp585_2015_2100_avg.csv")) %>% 
  rename(date = dayofyear, pcp = pr)

## daily totals
historical_tot = read.csv(here("climate_change_data","CNRM_daily_tot","oahu_historical_1850_2014_total.csv")) %>%  
  rename(date = dayofyear, pcp = pr) 
ssp126_tot = read.csv(here("climate_change_data","CNRM_daily_tot","oahu_ssp126_2015_2100_total.csv")) %>% 
  rename(date = dayofyear, pcp = pr) 
ssp245_tot = read.csv(here("climate_change_data","CNRM_daily_tot","oahu_ssp245_2015_2100_total.csv")) %>% 
  rename(date = dayofyear, pcp = pr) 
ssp370_tot = read.csv(here("climate_change_data","CNRM_daily_tot","oahu_ssp370_2015_2100_total.csv")) %>% 
  rename(date = dayofyear, pcp = pr)
ssp585_tot = read.csv(here("climate_change_data","CNRM_daily_tot","oahu_ssp585_2015_2100_total.csv")) %>% 
  rename(date = dayofyear, pcp = pr)

## convert date column to year, month, day components using luberdate package
historical_avg$date <- ymd(historical_avg$date)
ssp126_avg$date <- ymd(ssp126_avg$date)
ssp245_avg$date <- ymd(ssp245_avg$date)
ssp370_avg$date <- ymd(ssp370_avg$date)
ssp585_avg$date <- ymd(ssp585_avg$date)

historical_tot$date <- ymd(historical_tot$date)
ssp126_tot$date <- ymd(ssp126_tot$date)
ssp245_tot$date <- ymd(ssp245_tot$date)
ssp370_tot$date <- ymd(ssp370_tot$date)
ssp585_tot$date <- ymd(ssp585_tot$date)
```

## explore daily avg pcp data frame exported from python
```{r}
daily_avg_proj <- 
  ggplot() +
  geom_line(data = ssp126_avg, aes(x=date, y=pcp, color="orange"), size = 0.5) + 
  geom_line(data = ssp245_avg, aes(x=date, y=pcp, color="red"), size = 0.5) +
  geom_line(data = ssp370_avg, aes(x=date, y=pcp, color = "blue"), size = 0.5) +
  geom_line(data = ssp585_avg, aes(x=date, y=pcp, color = "seagreen"), size = 0.5) + 
  theme_minimal() + 
  scale_y_continuous(breaks=seq(0,7,1)) +
  scale_x_date(date_breaks = "10 years", date_labels = "%Y") +
  ggtitle("Oahu, HI Daily Avg Precipitation, CNRM-ESM2-1: 2015-2100") +
   labs(x = "Year",
       y = "Precipitation (in)") +
  scale_color_manual(values = c("orange", "red", "blue", "seagreen"), labels = c("spp126", "ssp245", "ssp370", "ssp585"))

daily_avg_proj

```

## wrangle data to have seperate yr, month, day columns
```{r}
# add seperate columns for year, month and day for avg daily data 
historical_avg <- historical_avg %>% 
  mutate(year = year(date)) %>% 
  mutate(month = month(date)) %>% 
  mutate(day = day(date))

ssp126_avg <- ssp126_avg %>% 
  mutate(year = year(date)) %>% 
  mutate(month = month(date)) %>% 
  mutate(day = day(date))

ssp245_avg <- ssp245_avg %>% 
  mutate(year = year(date)) %>% 
  mutate(month = month(date)) %>% 
  mutate(day = day(date))

ssp370_avg <- ssp370_avg %>% 
  mutate(year = year(date)) %>% 
  mutate(month = month(date)) %>% 
  mutate(day = day(date))

ssp585_avg <- ssp585_avg  %>% 
  mutate(year = year(date)) %>% 
  mutate(month = month(date)) %>% 
  mutate(day = day(date))


# group by yearly sum of pcp 
historical_yearly_avg <- historical_avg %>% 
  group_by(year) %>% 
  mutate(sum = sum(pcp)) %>% 
  select(year, sum) %>% 
  distinct()

historical_yearly_avg$year <- as.numeric(historical_yearly_avg$year) 


# group by yearly sum of pcp 
ssp126_yearly_avg <- ssp126_avg %>% 
  group_by(year) %>% 
  mutate(sum = sum(pcp)) %>% 
  select(year, sum) %>% 
  distinct()

ssp126_yearly_avg$year <- as.numeric(ssp126_yearly_avg$year) 

# group by yearly sum of pcp 
ssp245_yearly_avg <- ssp245_avg %>% 
  group_by(year) %>% 
  mutate(sum = sum(pcp)) %>% 
  select(year, sum) %>% 
  distinct()

ssp245_yearly_avg$year <- as.numeric(ssp245_yearly_avg$year) 

# group by yearly sum of pcp 
ssp370_yearly_avg <- ssp370_avg %>% 
  group_by(year) %>% 
  mutate(sum = sum(pcp)) %>% 
  select(year, sum) %>% 
  distinct()

ssp370_yearly_avg$year <- as.numeric(ssp370_yearly_avg$year) 

# group by yearly sum of pcp 
ssp585_yearly_avg <- ssp585_avg %>% 
  group_by(year) %>% 
  mutate(sum = sum(pcp)) %>% 
  select(year, sum) %>% 
  distinct()

ssp585_yearly_avg$year <- as.numeric(ssp585_yearly_avg$year) 


# group by month (and associated year) sum of pcp
historical_monthly_avg <- historical_avg %>% 
  group_by(month,year) %>% 
  mutate(sum = sum(pcp)) %>% 
  select(month, year, sum) %>% 
  distinct()

historical_monthly_avg$date <- format(as.Date(paste(historical_monthly_avg$year, historical_monthly_avg$month, 01), "%Y %m %d"), "%m-%Y")

historical_monthly_avg$date <- my(historical_monthly_avg$date)


# group by month (and associated year) sum of pcp
ssp126_monthly_avg <- ssp126_avg %>% 
  group_by(month,year) %>% 
  mutate(sum = sum(pcp)) %>% 
  select(month, year, sum) %>% 
  distinct()

ssp126_monthly_avg$date <- format(as.Date(paste(ssp126_monthly_avg$year, ssp126_monthly_avg$month, 01), "%Y %m %d"), "%m-%Y")

ssp126_monthly_avg$date <- my(ssp126_monthly_avg$date)


# group by month (and associated year) sum of pcp
ssp245_monthly_avg <- ssp245_avg %>% 
  group_by(month,year) %>% 
  mutate(sum = sum(pcp)) %>% 
  select(month, year, sum) %>% 
  distinct()

ssp245_monthly_avg$date <- format(as.Date(paste(ssp245_monthly_avg$year, ssp245_monthly_avg$month, 01), "%Y %m %d"), "%m-%Y")

ssp245_monthly_avg$date <- my(ssp245_monthly_avg$date)

# group by yearly sum of daily avg pcp 
ssp370_monthly_avg <- ssp370_avg %>% 
  group_by(month,year) %>% 
  mutate(sum = sum(pcp)) %>% 
  select(month, year, sum) %>% 
  distinct() 
  
ssp370_monthly_avg$date <- format(as.Date(paste(ssp370_monthly_avg$year, ssp370_monthly_avg$month, 01), "%Y %m %d"), "%m-%Y")

ssp370_monthly_avg$date <- my(ssp370_monthly_avg$date)

# group by yearly sum of daily avg pcp 
ssp585_monthly_avg <- ssp585_avg %>% 
  group_by(month,year) %>% 
  mutate(sum = sum(pcp)) %>% 
  select(month, year, sum) %>% 
  distinct() 
  
ssp585_monthly_avg$date <- format(as.Date(paste(ssp585_monthly_avg$year, ssp585_monthly_avg$month, 01), "%Y %m %d"), "%m-%Y")

ssp585_monthly_avg$date <- my(ssp585_monthly_avg$date)

```


## yearly projected precipiataion analysis 
```{r}
## all three ssps 
yearly_avg_proj <- 
  ggplot() +
  geom_line(data = ssp126_yearly_avg, aes(x=year, y=sum, color="orange"), size = 0.5) + 
  geom_line(data = ssp245_yearly_avg, aes(x=year, y=sum, color="red"), size = 0.5) +
  geom_line(data = ssp370_yearly_avg, aes(x=year, y=sum, color = "blue"), size = 0.5) +
  geom_line(data = ssp585_yearly_avg, aes(x=year, y=sum, color = "seagreen"), size = 0.5) +
  theme_minimal() + 
  scale_y_continuous(breaks=seq(0,50,10)) +
  scale_x_continuous(breaks=seq(2015,2100,10)) + 
  # scale_x_date(date_breaks = "5 years", date_labels = "%Y") +
  ggtitle("Oahu, HI Yearly Sum of Daily Avg Precipitation, CNRM-ESM2-1: 2015-2100") + 
  scale_color_manual(values = c("orange", "red", "blue", "seagreen"), labels = c("ssp126", "ssp245", "ssp370", "ssp585")) +
  labs(x = "Year",
       y = "Precipitation (in)") +
  geom_smooth(data = ssp126_yearly_avg, aes(x=year, y=sum), method='lm', formula = y~x, color = "orange") + 
  geom_smooth(data = ssp245_yearly_avg, aes(x=year, y=sum), method='lm', formula = y~x, color = "red") + 
  geom_smooth(data = ssp370_yearly_avg, aes(x=year, y=sum), method='lm', formula = y~x, color = "blue") + 
  geom_smooth(data = ssp585_yearly_avg, aes(x=year, y=sum), method='lm', formula = y~x, color = "seagreen") + 
  stat_regline_equation(label.y = 45) +
  stat_cor(label.y = 40)

yearly_avg_proj

## historical 
historical_yearly_avg_plot <- 
  ggplot(data = historical_yearly_avg, aes(x=year, y=sum)) +
  geom_line(size = 0.5,color="purple") +
  theme_minimal() + 
  scale_y_continuous(breaks=seq(0,50,10)) +
  scale_x_continuous(breaks=seq(1850,2015,10)) + 
  ggtitle("Oahu, HI Yearly Sum of Daily Avg Precipitation, historical, CNRM-ESM2-1: 1850-2014") + 
  geom_smooth(method='lm', formula = y~x, color = "purple") +
  stat_regline_equation(label.y = 45) +
  stat_cor(label.y = 40) +
   labs(x = "Year",
       y = "Precipitation (in)")

historical_yearly_avg_plot


## ssp126
ssp126_yearly_avg_plot <- 
  ggplot(data = ssp126_yearly_avg, aes(x=year, y=sum)) +
  geom_line(size = 0.5,color="orange") +
  theme_minimal() + 
  scale_y_continuous(breaks=seq(0,50,10)) +
  scale_x_continuous(breaks=seq(2015,2100,10)) + 
  ggtitle("Oahu, HI Yearly Sum of Daily Avg Precipitation,ssp126, CNRM-ESM2-1: 2015-2100") + 
  geom_smooth(method='lm', formula = y~x, color = "orange") +
  stat_regline_equation(label.y = 45) +
  stat_cor(label.y = 40) +
   labs(x = "Year",
       y = "Precipitation (in)")

ssp126_yearly_avg_plot


## ssp245
ssp245_yearly_avg_plot <- 
  ggplot(data = ssp245_yearly_avg, aes(x=year, y=sum)) +
  geom_line(size = 0.5,color="red") +
  theme_minimal() + 
  scale_y_continuous(breaks=seq(0,50,10)) +
  scale_x_continuous(breaks=seq(2015,2100,10)) + 
  ggtitle("Oahu, HI Yearly Sum of Daily Avg Precipitation,ssp245, CNRM-ESM2-1: 2015-2100") + 
  geom_smooth(method='lm', formula = y~x, color = "red") +
  stat_regline_equation(label.y = 45) +
  stat_cor(label.y = 40) +
   labs(x = "Year",
       y = "Precipitation (in)")

ssp245_yearly_avg_plot
  
## ssp370
ssp370_yearly_avg_plot <- 
  ggplot(data = ssp370_yearly_avg, aes(x=year, y=sum)) +
  geom_line(size = 0.5,color="blue") +
  theme_minimal() + 
  scale_y_continuous(breaks=seq(0,50,10)) +
  scale_x_continuous(breaks=seq(2015,2100,10)) + 
  ggtitle("Oahu, HI Yearly Sum of Daily Avg Precipitation, ssp370, CNRM-ESM2-1: 2015-2100") + 
  geom_smooth(method='lm', formula = y~x, color = "blue") +
  stat_regline_equation(label.y = 45) +
  stat_cor(label.y = 40) +
   labs(x = "Year",
       y = "Precipitation (in)")

ssp370_yearly_avg_plot

## ssp585
ssp585_yearly_avg_plot <- 
  ggplot(data = ssp585_yearly_avg, aes(x=year, y=sum)) +
  geom_line(size = 0.5,color="seagreen") +
  theme_minimal() + 
  scale_y_continuous(breaks=seq(0,50,10)) +
  scale_x_continuous(breaks=seq(2015,2100,10)) + 
  ggtitle("Oahu, HI Yearly Sum of Daily Avg Precipitation, ssp585, CNRM-ESM2-1: 2015-2100") + 
  geom_smooth(method='lm', formula = y~x, color = "seagreen") +
  stat_regline_equation(label.y = 45) +
  stat_cor(label.y = 40) +
   labs(x = "Year",
       y = "Precipitation (in)")

ssp585_yearly_avg_plot
```

## month by year projected precipiation analysis
```{r}
## historical monthly 
historical_monthly_avg_plot <- 
  ggplot(data = historical_monthly_avg, aes(x=date, y=sum)) +
  geom_line(size = 0.5,color="purple") +
  theme_minimal() + 
  scale_y_continuous(breaks=seq(0,20,5)) +
  #scale_x_continuous(breaks=seq(2015,2100,10)) + 
  ggtitle("Oahu, HI Monthly Sum of Daily Avg Precipitation, historical, CNRM-ESM2-1: 2015-2100") + 
  geom_smooth(method='lm', formula = y~x, color = "purple") +
  stat_regline_equation(label.y = 16) +
  stat_cor(label.y = 15) +
   labs(x = "Year",
       y = "Precipitation (in)")

historical_monthly_avg_plot


## ssp126 monthly 
ssp126_monthly_avg_plot <- 
  ggplot(data = ssp126_monthly_avg, aes(x=date, y=sum)) +
  geom_line(size = 0.5,color="orange") +
  theme_minimal() + 
  scale_y_continuous(breaks=seq(0,20,5)) +
  #scale_x_continuous(breaks=seq(2015,2100,10)) + 
  ggtitle("Oahu, HI Monthly Sum of Daily Avg Precipitation, ssp126, CNRM-ESM2-1: 2015-2100") + 
  geom_smooth(method='lm', formula = y~x, color = "orange") +
  stat_regline_equation(label.y = 16) +
  stat_cor(label.y = 15) +
   labs(x = "Year",
       y = "Precipitation (in)")

ssp126_monthly_avg_plot

## ssp245 monthly 
ssp245_monthly_avg_plot <- 
  ggplot(data = ssp245_monthly_avg, aes(x=date, y=sum)) +
  geom_line(size = 0.5,color="red") +
  theme_minimal() + 
  scale_y_continuous(breaks=seq(0,20,5)) +
  #scale_x_continuous(breaks=seq(2015,2100,10)) + 
  ggtitle("Oahu, HI Monthly Sum of Daily Avg Precipitation, ssp245, CNRM-ESM2-1: 2015-2100") + 
  geom_smooth(method='lm', formula = y~x, color = "red") +
  stat_regline_equation(label.y = 16) +
  stat_cor(label.y = 15) +
   labs(x = "Year",
       y = "Precipitation (in)")

ssp245_monthly_avg_plot

## ssp370 monthly
ssp370_monthly_avg_plot <- 
  ggplot(data = ssp370_monthly_avg, aes(x=date, y=sum)) +
  geom_line(size = 0.5,color="blue") +
  theme_minimal() + 
  scale_y_continuous(breaks=seq(0,20,5)) +
  #scale_x_continuous(breaks=seq(2015,2100,10)) + 
  ggtitle("Oahu, HI Monthly Sum of Daily Avg Precipitation, ssp370, CNRM-ESM2-1: 2015-2100") + 
  geom_smooth(method='lm', formula = y~x, color = "blue") +
  stat_regline_equation(label.y = 16) +
  stat_cor(label.y = 15) +
   labs(x = "Year",
       y = "Precipitation (in)")

ssp370_monthly_avg_plot

## ssp585 monthly
ssp585_monthly_avg_plot <- 
  ggplot(data = ssp585_monthly_avg, aes(x=date, y=sum)) +
  geom_line(size = 0.5,color="seagreen") +
  theme_minimal() + 
  scale_y_continuous(breaks=seq(0,20,5)) +
  #scale_x_continuous(breaks=seq(2015,2100,10)) + 
  ggtitle("Oahu, HI Monthly Sum of Daily Avg Precipitation, ssp585, CNRM-ESM2-1: 2015-2100") + 
  geom_smooth(method='lm', formula = y~x, color = "seagreen") +
  stat_regline_equation(label.y = 16) +
  stat_cor(label.y = 15) + 
   labs(x = "Year",
       y = "Precipitation (in)")

ssp585_monthly_avg_plot
```
## monthly combined projected precipation analysis 
```{r}
monthly_avg_proj <- 
  ggplot() +
  geom_line(data = ssp126_monthly_avg, aes(x=date, y=sum, color= "orange"), size = 0.5) + 
  geom_line(data = ssp245_monthly_avg, aes(x=date, y=sum, color= "red"), size = 0.5) +
  geom_line(data = ssp370_monthly_avg, aes(x=date, y=sum, color = "blue"), size = 0.5) +
  geom_line(data = ssp585_monthly_avg, aes(x=date, y=sum, color = "seagreen"), size = 0.5) +
  theme_minimal() + 
  scale_y_continuous(breaks=seq(0,20,5)) +
  #scale_x_continuous(breaks=seq(2015,2100,10)) + 
  # scale_x_date(date_breaks = "5 years", date_labels = "%Y") +
  ggtitle("Oahu, HI Monthly Sum of Daily Avg Precipitation, CNRM-ESM2-1: 2015-2100") + 
  scale_color_manual(values = c("orange", "red", "blue", "seagreen"), labels = c("ssp126", "ssp245", "ssp370", "ssp585")) +
  labs(x = "Year",
       y = "Precipitation (in)") 

monthly_avg_proj
```

## export graphs 
```{r}
# ggsave("img/daily_avg_proj.png", daily_avg_proj, height = 5, width = 7, dpi = 320)
# ggsave("img/yearly_avg_proj.png", yearly_avg_proj, height = 5, width = 7, dpi = 320)
# ggsave("img/monthly_avg_proj.png", monthly_avg_proj, height = 5, width = 7, dpi = 320)
# ggsave("img/ssp245_yearly_plot.png", ssp245_yearly_plot, height = 5, width = 7, dpi = 320)
# ggsave("img/ssp370_yearly_plot.png", ssp370_yearly_plot, height = 5, width = 7, dpi = 320)
# ggsave("img/ssp585_yearly_plot.png", ssp585_yearly_plot, height = 5, width = 7, dpi = 320)
# ggsave("img/ssp245_monthly_plot.png", ssp245_monthly_plot, height = 5, width = 7, dpi = 320)
# ggsave("img/ssp370_monthly_plot.png", ssp370_monthly_plot, height = 5, width = 7, dpi = 320)
# ggsave("img/ssp585_monthly_plot.png", ssp585_monthly_plot, height = 5, width = 7, dpi = 320)
```


## Daily total analysis 
```{r}
daily_tot_proj <- 
  ggplot() +
  geom_line(data = historical_tot, aes(x=date, y=pcp, color= "black"), size = 0.5) + 
  geom_line(data = ssp126_tot, aes(x=date, y=pcp, color="orange"), size = 0.5) + 
  geom_line(data = ssp245_tot, aes(x=date, y=pcp, color="red"), size = 0.5) +
  geom_line(data = ssp370_tot, aes(x=date, y=pcp, color = "blue"), size = 0.5) +
  geom_line(data = ssp585_tot, aes(x=date, y=pcp, color = "seagreen"), size = 0.5) + 
  theme_minimal() + 
  scale_y_continuous(breaks=seq(0,50,10)) +
  scale_x_date(date_breaks = "20 years", date_labels = "%Y") +
  ggtitle("Oahu, HI Daily Total Precipitation, CNRM-ESM2-1: 1850-2100") +
   labs(x = "Year",
       y = "Precipitation (in)") +
  scale_color_manual(values = c("black", "orange", "red", "blue", "seagreen", "purple"), 
                     labels = c("historical" , "spp126", "ssp245", "ssp370", "ssp585")) +
  theme(legend.title=element_blank())+
    theme(legend.position="bottom")

daily_tot_proj

```

## daily totals by month 
```{r}
# add seperate columns for year, month and day
historical_tot <- historical_tot %>% 
  mutate(year = year(date)) %>% 
  mutate(month = month(date)) %>% 
  mutate(day = day(date))

ssp126_tot <- ssp126_tot %>% 
  mutate(year = year(date)) %>% 
  mutate(month = month(date)) %>% 
  mutate(day = day(date))

ssp245_tot <- ssp245_tot %>% 
  mutate(year = year(date)) %>% 
  mutate(month = month(date)) %>% 
  mutate(day = day(date))

ssp370_tot <- ssp370_tot %>% 
  mutate(year = year(date)) %>% 
  mutate(month = month(date)) %>% 
  mutate(day = day(date))

ssp585_tot <- ssp585_tot  %>% 
  mutate(year = year(date)) %>% 
  mutate(month = month(date)) %>% 
  mutate(day = day(date))


# group by yearly sum of pcp 
historical_tot_yearly <- historical_tot %>% 
  group_by(year) %>% 
  mutate(sum = sum(pcp)) %>% 
  select(year, sum) %>% 
  distinct()

historical_tot_yearly$year <- as.numeric(historical_tot_yearly$year) 


# group by yearly sum of pcp 
ssp126_tot_yearly <- ssp126_tot %>% 
  group_by(year) %>% 
  mutate(sum = sum(pcp)) %>% 
  select(year, sum) %>% 
  distinct()

ssp126_tot_yearly$year <- as.numeric(ssp126_tot_yearly$year) 

# group by yearly sum of pcp 
ssp245_tot_yearly <- ssp245_tot %>% 
  group_by(year) %>% 
  mutate(sum = sum(pcp)) %>% 
  select(year, sum) %>% 
  distinct()

ssp245_tot_yearly$year <- as.numeric(ssp245_tot_yearly$year) 

# group by yearly sum of pcp 
ssp370_tot_yearly <- ssp370_tot %>% 
  group_by(year) %>% 
  mutate(sum = sum(pcp)) %>% 
  select(year, sum) %>% 
  distinct()

ssp370_tot_yearly$year <- as.numeric(ssp370_tot_yearly$year) 

# group by yearly sum of pcp 
ssp585_tot_yearly <- ssp585_tot %>% 
  group_by(year) %>% 
  mutate(sum = sum(pcp)) %>% 
  select(year, sum) %>% 
  distinct()

ssp585_tot_yearly$year <- as.numeric(ssp585_tot_yearly$year) 


# group by month (and associated year) sum of pcp
historical_tot_monthly <- historical_tot %>% 
  group_by(month,year) %>% 
  mutate(sum = sum(pcp)) %>% 
  select(month, year, sum) %>% 
  distinct()

historical_tot_monthly$date <- format(as.Date(paste(historical_tot_monthly$year, historical_tot_monthly$month, 01), "%Y %m %d"), "%m-%Y")

historical_tot_monthly$date <- my(historical_tot_monthly$date)


# group by month (and associated year) sum of pcp
ssp126_tot_monthly <- ssp126_tot %>% 
  group_by(month,year) %>% 
  mutate(sum = sum(pcp)) %>% 
  select(month, year, sum) %>% 
  distinct()

ssp126_tot_monthly$date <- format(as.Date(paste(ssp126_tot_monthly$year, ssp126_tot_monthly$month, 01), "%Y %m %d"), "%m-%Y")

ssp126_tot_monthly$date <- my(ssp126_tot_monthly$date)


# group by month (and associated year) sum of pcp
ssp245_tot_monthly <- ssp245_tot %>% 
  group_by(month,year) %>% 
  mutate(sum = sum(pcp)) %>% 
  select(month, year, sum) %>% 
  distinct()

ssp245_tot_monthly$date <- format(as.Date(paste(ssp245_tot_monthly$year, ssp245_tot_monthly$month, 01), "%Y %m %d"), "%m-%Y")

ssp245_tot_monthly$date <- my(ssp245_tot_monthly$date)

# group by yearly sum of daily avg pcp 
ssp370_tot_monthly <- ssp370_tot %>% 
  group_by(month,year) %>% 
  mutate(sum = sum(pcp)) %>% 
  select(month, year, sum) %>% 
  distinct() 
  
ssp370_tot_monthly$date <- format(as.Date(paste(ssp370_tot_monthly$year, ssp370_tot_monthly$month, 01), "%Y %m %d"), "%m-%Y")

ssp370_tot_monthly$date <- my(ssp370_tot_monthly$date)

# group by yearly sum of daily avg pcp 
ssp585_tot_monthly <- ssp585_tot %>% 
  group_by(month,year) %>% 
  mutate(sum = sum(pcp)) %>% 
  select(month, year, sum) %>% 
  distinct() 
  
ssp585_tot_monthly$date <- format(as.Date(paste(ssp585_tot_monthly$year, ssp585_tot_monthly$month, 01), "%Y %m %d"), "%m-%Y")

ssp585_tot_monthly$date <- my(ssp585_tot_monthly$date)
```

## yearly total for all ssps and historical
```{r}
## all three ssps plotted below
yearly_sum_proj <- 
  ggplot() +
  geom_line(data = historical_tot_yearly, aes(x=year, y=sum, color="black"), size = 0.5) +
  geom_line(data = ssp126_tot_yearly, aes(x=year, y=sum, color="orange"), size = 0.5) + 
  geom_line(data = ssp245_tot_yearly, aes(x=year, y=sum, color="red"), size = 0.5) +
  geom_line(data = ssp370_tot_yearly, aes(x=year, y=sum, color = "blue"), size = 0.5) +
  geom_line(data = ssp585_tot_yearly, aes(x=year, y=sum, color = "seagreen"), size = 0.5) +
  theme_minimal() + 
  scale_y_continuous(breaks=seq(0,450,50)) +
  scale_x_continuous(breaks=seq(1850,2100,50)) + 
  ggtitle("Oahu, HI Yearly Total Precipitation, CNRM-ESM2-1: 1850-2100") + 
  scale_color_manual(values = c("black", "orange", "red", "blue", "seagreen"), labels = c("historical", "ssp126", "ssp245", "ssp370", "ssp585")) +
  labs(x = "Year",
       y = "Precipitation (in)") +
  geom_smooth(data = historical_tot_yearly, aes(x=year, y=sum), method='lm', formula = y~x, color = "black") +
  geom_smooth(data = ssp126_tot_yearly, aes(x=year, y=sum), method='lm', formula = y~x, color = "orange") + 
  geom_smooth(data = ssp245_tot_yearly, aes(x=year, y=sum), method='lm', formula = y~x, color = "red") + 
  geom_smooth(data = ssp370_tot_yearly, aes(x=year, y=sum), method='lm', formula = y~x, color = "blue") + 
  geom_smooth(data = ssp585_tot_yearly, aes(x=year, y=sum), method='lm', formula = y~x, color = "seagreen") +
  theme(legend.title=element_blank())+
    theme(legend.position="bottom")

yearly_sum_proj
```


## monthly total for all ssps and historical
```{r}
# all three ssps 
monthly_sum_proj <- 
  ggplot() +
  geom_line(data = historical_tot_monthly, aes(x=date, y=sum, color="black"), size = 0.5) +
  geom_line(data = ssp126_tot_monthly, aes(x=date, y=sum, color="orange"), size = 0.5) + 
  geom_line(data = ssp245_tot_monthly, aes(x=date, y=sum, color="red"), size = 0.5) +
  geom_line(data = ssp370_tot_monthly, aes(x=date, y=sum, color = "blue"), size = 0.5) +
  geom_line(data = ssp585_tot_monthly, aes(x=date, y=sum, color = "seagreen"), size = 0.5) +
  theme_minimal() + 
  scale_y_continuous(breaks=seq(0,300,50)) +
 scale_x_date(date_breaks = "50 years", date_labels = "%Y") +
  ggtitle("Oahu, HI Monthly Total Precipitation, CNRM-ESM2-1: 1850-2100") + 
  scale_color_manual(values = c("black", "orange", "red", "blue", "seagreen"), labels = c("historical", "ssp126", "ssp245", "ssp370", "ssp585")) +
  labs(x = "Year",
       y = "Precipitation (in)") +
  # geom_smooth(data = historical_tot_monthly, aes(x=date, y=sum), method='lm', formula = y~x, color = "black") +
  # geom_smooth(data = ssp126_tot_monthly, aes(x=date, y=sum), method='lm', formula = y~x, color = "orange") + 
  # geom_smooth(data = ssp245_tot_monthly, aes(x=date, y=sum), method='lm', formula = y~x, color = "red") + 
  # geom_smooth(data = ssp370_tot_monthly, aes(x=date, y=sum), method='lm', formula = y~x, color = "blue") + 
  # geom_smooth(data = ssp585_tot_monthly, aes(x=date, y=sum), method='lm', formula = y~x, color = "seagreen") +
  theme(legend.title=element_blank())+
    theme(legend.position="bottom")

monthly_sum_proj
```
## separate historical total precip based on 30 yr time periods
```{r}
cnrm_historical_tot_1984_2014 <- historical_tot %>% 
  filter(year %in% (1984:2014))
  
## separate projections based on 2020 to 2070
cnrm_ssp126_tot_2020_2050 <- ssp126_tot %>% 
  filter(year %in% (2020:2050))

cnrm_ssp245_tot_2020_2050 <- ssp245_tot %>% 
  filter(year %in% (2020:2050))

cnrm_ssp370_tot_2020_2050 <- ssp370_tot %>% 
  filter(year %in% (2020:2050))

cnrm_ssp585_tot_2020_2050 <- ssp585_tot %>% 
  filter(year %in% (2020:2050))


cnrm_daily_tot_period <- 
  ggplot() +
  geom_line(data = cnrm_historical_tot_1984_2014, aes(x=date, y=pcp, color="black"), size = 0.5) + 
  geom_line(data = cnrm_ssp126_tot_2020_2050, aes(x=date, y=pcp, color="orange"), size = 0.5) + 
  geom_line(data = cnrm_ssp245_tot_2020_2050, aes(x=date, y=pcp, color="red"), size = 0.5) +
  geom_line(data = cnrm_ssp370_tot_2020_2050, aes(x=date, y=pcp, color = "blue"), size = 0.5) +
  geom_line(data = cnrm_ssp585_tot_2020_2050, aes(x=date, y=pcp, color = "seagreen"), size = 0.5) + 
  theme_minimal() + 
  scale_y_continuous(breaks=seq(0,100,10), limits = c(0, 100)) +
  scale_x_date(date_breaks = "3 years", date_labels = "%Y") +
  ggtitle("Oahu, HI Daily Total Precipitation, CNRM-ESM2-1: 1984-2050") +
   labs(x = "Year",
       y = "Precipitation (in)") +
  scale_color_manual(values = c("black", "orange", "red", "blue", "seagreen"), labels = c("historical", "spp126", "ssp245", "ssp370", "ssp585")) +
   theme(legend.title=element_blank())+
    theme(legend.position="bottom") +
    theme(text = element_text(size=14))

cnrm_daily_tot_period

#ggsave('cnrm_daily_tot_period.png', cnrm_daily_tot_period, width = 12, height = 8, units = "in")

```

## historical 50 year period graph
```{r}
cnrm_daily_tot_period_hist <- 
  ggplot() +
  geom_line(data = cnrm_historical_tot_1984_2014, aes(x=date, y=pcp, color="black"), size = 0.5) +  
  theme_minimal() + 
  scale_y_continuous(breaks=seq(0,50,10)) +
  scale_x_date(date_breaks = "3.5 years", date_labels = "%Y") +
  ggtitle("Oahu, HI Daily Total Precipitation, CNRM-ESM2-1, historical: 1984-2014 ") +
   labs(x = "Year",
       y = "Precipitation (in)") +
  scale_color_manual(values = c("black"), labels = c("historical")) + 
 theme(legend.title=element_blank()) +
    theme(legend.position="bottom")

cnrm_daily_tot_period_hist
```

## ratio of daily total precipitation in the future vs. historical periods
```{r}
## merge the data frames together 
daily_tot_prec_merge1 <- cnrm_ssp126_tot_2020_2050 %>% right_join(cnrm_ssp245_tot_2020_2050, by = c("date")) 
daily_tot_prec_merge2 <- daily_tot_prec_merge1 %>%  right_join(cnrm_ssp370_tot_2020_2050, by = c("date"))
daily_tot_prec_merge_fin <- daily_tot_prec_merge2 %>% right_join(cnrm_ssp585_tot_2020_2050,by = c("date")) %>% 
  select(date, year.x, month.x, day.x, pcp.x, pcp.y, pcp.x.x, pcp.y.y) %>% 
  rename(year = year.x,
         month = month.x,
         day = day.x,
         ssp126_pcp=pcp.x,
         ssp245_pcp=pcp.y,
         ssp370_pcp = pcp.x.x ,
         ssp585_pcp = pcp.y.y )

historical_tot_1984_2014_rename <- cnrm_historical_tot_1984_2014 %>% 
  rename(hist_date = date,
         hist_pcp = pcp,
         hist_year = year,
         hist_month = month,
         hist_day = day)

merge_his_proj <- cbind(daily_tot_prec_merge_fin, historical_tot_1984_2014_rename) 

ratio_df <- merge_his_proj %>% 
  mutate(ssp126_hist = ssp126_pcp/hist_pcp,
         ssp245_hist = ssp245_pcp/hist_pcp,
         ssp370_hist = ssp370_pcp/hist_pcp,
         ssp585_hist = ssp585_pcp/hist_pcp) %>% 
  select(date, ssp126_hist,ssp245_hist, ssp370_hist, ssp585_hist) %>% 
  pivot_longer(cols = ssp126_hist:ssp585_hist,
               names_to = c("ssp"),
               values_to = "fut_hist_ratio")

ggplot(data = ratio_df, aes(x=date, y=fut_hist_ratio,  group=ssp)) +
  geom_boxplot() +
  facet_wrap(~ssp)
  scale_y_continuous(limits = c(0,0.0e+00))
```


## December 19, 2010 = One Wailupe storm that we will use to do climate analysis with. Scrapping Kuliouou Oct 11th, 2009 storm, given we're having issues getting accurate SWMM output that correlated to the stream gage data
```{r}
merge_his_proj_filt_dec19 <- merge_his_proj %>% 
  filter(month == "12") %>% 
  filter(day == "19") %>% 
  group_by(month,day) %>% 
  mutate(ssp126_pcp_avg = mean(ssp126_pcp),
         ssp245_pcp_avg = mean(ssp245_pcp),
         ssp370_pcp_avg = mean(ssp370_pcp),
         ssp585_pcp_avg = mean(ssp585_pcp),
         hist_pcp_avg = mean(hist_pcp)) %>% 
  select(month,day,ssp126_pcp_avg,ssp245_pcp_avg,ssp370_pcp_avg, ssp585_pcp_avg, hist_pcp_avg ) %>% 
  distinct()


merge_his_proj_filt_dec19 <- merge_his_proj_filt_dec19 %>% 
  mutate(ssp126_hist = ssp126_pcp_avg/hist_pcp_avg,
         ssp245_hist = ssp245_pcp_avg/hist_pcp_avg,
         ssp370_hist = ssp370_pcp_avg/hist_pcp_avg,
         ssp585_hist = ssp585_pcp_avg/hist_pcp_avg) %>% 
  select(month, day, ssp126_hist,ssp245_hist, ssp370_hist, ssp585_hist) %>% 
  pivot_longer(cols = ssp126_hist:ssp585_hist,
               names_to = c("ssp"),
               values_to = "fut_hist_ratio")

proj_hist_dec19_cnrm <- ggplot(data = merge_his_proj_filt_dec19, aes(x=ssp, y=fut_hist_ratio)) +
  geom_point(size=2) +
  theme_minimal() +
  labs(x="Shared Socioeconomic Pathways (ssp)", 
       y="Future Projection/Historical Daily Total Average Ratio (Dec 19)", 
       caption = "Climate Model: CNRM-ESM2-1") +
    theme(text = element_text(size=15)) +
  scale_x_discrete(labels=c("ssp126_hist" = "ssp 1-2.6", 
                            "ssp245_hist" = "ssp 2-4.5",
                              "ssp370_hist" = "ssp 3-7.0",
                            "ssp585_hist" = "ssp 5-8.5")) +
  scale_y_continuous(limits = c(0,3.5))

proj_hist_dec19_cnrm

 # ggsave('proj_hist_dec19_cnrm.png', proj_hist_dec19_cnrm, width = 12, height = 10, units = "in")

```

## March 14, 2009 Wailupe storm that we will use to do Climate Change analysis with. Here we are taking up the ratio of projected daily total precip avg from 2020-2050 / historical daily total precip avg from 1984-2014. 
```{r}
merge_his_proj_filt_mar14 <- merge_his_proj %>% 
  filter(month == "3") %>% 
  filter(day == "14") %>% 
  group_by(month,day) %>% 
  mutate(ssp126_pcp_avg = mean(ssp126_pcp),
         ssp245_pcp_avg = mean(ssp245_pcp),
         ssp370_pcp_avg = mean(ssp370_pcp),
         ssp585_pcp_avg = mean(ssp585_pcp),
         hist_pcp_avg = mean(hist_pcp)) %>% 
  select(month,day,ssp126_pcp_avg,ssp245_pcp_avg,ssp370_pcp_avg, ssp585_pcp_avg, hist_pcp_avg ) %>% 
  distinct()


merge_his_proj_filt_mar14 <- merge_his_proj_filt_mar14 %>% 
  mutate(ssp126_hist = ssp126_pcp_avg/hist_pcp_avg,
         ssp245_hist = ssp245_pcp_avg/hist_pcp_avg,
         ssp370_hist = ssp370_pcp_avg/hist_pcp_avg,
         ssp585_hist = ssp585_pcp_avg/hist_pcp_avg) %>% 
  select(month, day, ssp126_hist,ssp245_hist, ssp370_hist, ssp585_hist) %>% 
  pivot_longer(cols = ssp126_hist:ssp585_hist,
               names_to = c("ssp"),
               values_to = "fut_hist_ratio")

proj_hist_mar14_cnrm <- ggplot(data = merge_his_proj_filt_mar14, aes(x=ssp, y=fut_hist_ratio)) +
  geom_point(size=2) +
  theme_minimal() +
  labs(x="Shared Socioeconomic Pathways (ssp)", 
       y="Future Projection/Historical Daily Total Average Ratio (Mar 14)", 
       caption = "Climate Model: CNRM-ESM2-1") +
    theme(text = element_text(size=15)) +
  scale_x_discrete(labels=c("ssp126_hist" = "ssp 1-2.6", 
                            "ssp245_hist" = "ssp 2-4.5",
                              "ssp370_hist" = "ssp 3-7.0",
                            "ssp585_hist" = "ssp 5-8.5")) +
  scale_y_continuous(limits = c(0,3.5))

proj_hist_mar14_cnrm

 # ggsave('proj_hist_mar14_cnrm.png', proj_hist_mar14_cnrm, width = 12, height = 10, units = "in")

```

