---
title: "Climate Change Projections: IPSL-CM6A-LR"
author: "Elmera Azadpour"
date: "1/14/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(lubridate)
library(ggpubr)
library(zoo)
```

## read in daily total climate projection and historical data exported from jupyter notebook, code can be found here: https://github.com/aloha-aina/Aloha-Aina-Master 
```{r}
## Note: these data are from climate model IPSL-CM6A-LR (France)
## daily totals
ipsl_historical_tot = read.csv(here("climate_change_data", "IPSL_daily_tot","oahu_historical_1850_2014_total.csv")) %>%  
  rename(date = dayofyear, pcp = pr) 
ipsl_ssp126_tot = read.csv(here("climate_change_data","IPSL_daily_tot","oahu_ssp126_2015_2100_total.csv")) %>% 
  rename(date = dayofyear, pcp = pr) 
ipsl_ssp245_tot = read.csv(here("climate_change_data","IPSL_daily_tot","oahu_ssp245_2015_2100_total.csv")) %>% 
  rename(date = dayofyear, pcp = pr) 
ipsl_ssp370_tot = read.csv(here("climate_change_data","IPSL_daily_tot","oahu_ssp370_2015_2100_total.csv")) %>% 
  rename(date = dayofyear, pcp = pr)
ipsl_ssp585_tot = read.csv(here("climate_change_data","IPSL_daily_tot","oahu_ssp585_2015_2100_total.csv")) %>% 
  rename(date = dayofyear, pcp = pr)

ipsl_historical_tot$date <- ymd(ipsl_historical_tot$date)
ipsl_ssp126_tot$date <- ymd(ipsl_ssp126_tot$date)
ipsl_ssp245_tot$date <- ymd(ipsl_ssp245_tot$date)
ipsl_ssp370_tot$date <- ymd(ipsl_ssp370_tot$date)
ipsl_ssp585_tot$date <- ymd(ipsl_ssp585_tot$date)
```

## Daily total visualization 
```{r}
ipsl_daily_tot_proj <- 
  ggplot() +
  geom_line(data = ipsl_historical_tot, aes(x=date, y=pcp, color= "black"), size = 0.5) + 
  geom_line(data = ipsl_ssp126_tot, aes(x=date, y=pcp, color="orange"), size = 0.5) + 
  geom_line(data = ipsl_ssp245_tot, aes(x=date, y=pcp, color="red"), size = 0.5) +
  geom_line(data = ipsl_ssp370_tot, aes(x=date, y=pcp, color = "blue"), size = 0.5) +
  geom_line(data = ipsl_ssp585_tot, aes(x=date, y=pcp, color = "seagreen"), size = 0.5) + 
  theme_minimal() + 
  scale_y_continuous(breaks=seq(0,150,10)) +
  scale_x_date(date_breaks = "20 years", date_labels = "%Y") +
  ggtitle("Oahu, HI Daily Total Precipitation, IPSL-CM6A-LR: 1850-2100") +
   labs(x = "Year",
       y = "Precipitation (in)") +
  scale_color_manual(values = c("black", "orange", "red", "blue", "seagreen", "purple"), 
                     labels = c("historical" , "spp126", "ssp245", "ssp370", "ssp585")) +
  theme(legend.title=element_blank())+
    theme(legend.position="bottom")

ipsl_daily_tot_proj

```

## daily totals by month 
```{r}
# add seperate columns for year, month and day
ipsl_historical_tot <- ipsl_historical_tot %>% 
  mutate(year = year(date)) %>% 
  mutate(month = month(date)) %>% 
  mutate(day = day(date))

ipsl_ssp126_tot <- ipsl_ssp126_tot %>% 
  mutate(year = year(date)) %>% 
  mutate(month = month(date)) %>% 
  mutate(day = day(date))

ipsl_ssp245_tot <- ipsl_ssp245_tot %>% 
  mutate(year = year(date)) %>% 
  mutate(month = month(date)) %>% 
  mutate(day = day(date))

ipsl_ssp370_tot <- ipsl_ssp370_tot %>% 
  mutate(year = year(date)) %>% 
  mutate(month = month(date)) %>% 
  mutate(day = day(date))

ipsl_ssp585_tot <- ipsl_ssp585_tot  %>% 
  mutate(year = year(date)) %>% 
  mutate(month = month(date)) %>% 
  mutate(day = day(date))


# group by yearly sum of pcp 
ipsl_historical_tot_yearly <- ipsl_historical_tot %>% 
  group_by(year) %>% 
  mutate(sum = sum(pcp)) %>% 
  select(year, sum) %>% 
  distinct()

ipsl_historical_tot_yearly$year <- as.numeric(ipsl_historical_tot_yearly$year) 


# group by yearly sum of pcp 
ipsl_ssp126_tot_yearly <- ipsl_ssp126_tot %>% 
  group_by(year) %>% 
  mutate(sum = sum(pcp)) %>% 
  select(year, sum) %>% 
  distinct()

ipsl_ssp126_tot_yearly$year <- as.numeric(ipsl_ssp126_tot_yearly$year) 

# group by yearly sum of pcp 
ipsl_ssp245_tot_yearly <- ipsl_ssp245_tot %>% 
  group_by(year) %>% 
  mutate(sum = sum(pcp)) %>% 
  select(year, sum) %>% 
  distinct()

ipsl_ssp245_tot_yearly$year <- as.numeric(ipsl_ssp245_tot_yearly$year) 

# group by yearly sum of pcp 
ipsl_ssp370_tot_yearly <- ipsl_ssp370_tot %>% 
  group_by(year) %>% 
  mutate(sum = sum(pcp)) %>% 
  select(year, sum) %>% 
  distinct()

ipsl_ssp370_tot_yearly$year <- as.numeric(ipsl_ssp370_tot_yearly$year) 

# group by yearly sum of pcp 
ipsl_ssp585_tot_yearly <- ipsl_ssp585_tot %>% 
  group_by(year) %>% 
  mutate(sum = sum(pcp)) %>% 
  select(year, sum) %>% 
  distinct()

ipsl_ssp585_tot_yearly$year <- as.numeric(ipsl_ssp585_tot_yearly$year) 


# group by month (and associated year) sum of pcp
ipsl_historical_tot_monthly <- ipsl_historical_tot %>% 
  group_by(month,year) %>% 
  mutate(sum = sum(pcp)) %>% 
  select(month, year, sum) %>% 
  distinct()

ipsl_historical_tot_monthly$date <- format(as.Date(paste(ipsl_historical_tot_monthly$year, ipsl_historical_tot_monthly$month, 01), "%Y %m %d"), "%m-%Y")

ipsl_historical_tot_monthly$date <- my(ipsl_historical_tot_monthly$date)


# group by month (and associated year) sum of pcp
ipsl_ssp126_tot_monthly <- ipsl_ssp126_tot %>% 
  group_by(month,year) %>% 
  mutate(sum = sum(pcp)) %>% 
  select(month, year, sum) %>% 
  distinct()

ipsl_ssp126_tot_monthly$date <- format(as.Date(paste(ipsl_ssp126_tot_monthly$year, ipsl_ssp126_tot_monthly$month, 01), "%Y %m %d"), "%m-%Y")

ipsl_ssp126_tot_monthly$date <- my(ipsl_ssp126_tot_monthly$date)


# group by month (and associated year) sum of pcp
ipsl_ssp245_tot_monthly <- ipsl_ssp245_tot %>% 
  group_by(month,year) %>% 
  mutate(sum = sum(pcp)) %>% 
  select(month, year, sum) %>% 
  distinct()

ipsl_ssp245_tot_monthly$date <- format(as.Date(paste(ipsl_ssp245_tot_monthly$year, ipsl_ssp245_tot_monthly$month, 01), "%Y %m %d"), "%m-%Y")

ipsl_ssp245_tot_monthly$date <- my(ipsl_ssp245_tot_monthly$date)

# group by yearly sum of daily avg pcp 
ipsl_ssp370_tot_monthly <- bcc_ssp370_tot %>% 
  group_by(month,year) %>% 
  mutate(sum = sum(pcp)) %>% 
  select(month, year, sum) %>% 
  distinct() 
  
ipsl_ssp370_tot_monthly$date <- format(as.Date(paste(ipsl_ssp370_tot_monthly$year, ipsl_ssp370_tot_monthly$month, 01), "%Y %m %d"), "%m-%Y")

ipsl_ssp370_tot_monthly$date <- my(ipsl_ssp370_tot_monthly$date)

# group by yearly sum of daily avg pcp 
ipsl_ssp585_tot_monthly <- ipsl_ssp585_tot %>% 
  group_by(month,year) %>% 
  mutate(sum = sum(pcp)) %>% 
  select(month, year, sum) %>% 
  distinct() 
  
ipsl_ssp585_tot_monthly$date <- format(as.Date(paste(ipsl_ssp585_tot_monthly$year, ipsl_ssp585_tot_monthly$month, 01), "%Y %m %d"), "%m-%Y")

ipsl_ssp585_tot_monthly$date <- my(ipsl_ssp585_tot_monthly$date)
```

## yearly total for all ssps and historical
```{r}
## all three ssps 
ipsl_yearly_sum_proj <- 
  ggplot() +
  geom_line(data = ipsl_historical_tot_yearly, aes(x=year, y=sum, color="black"), size = 0.5) +
  geom_line(data = ipsl_ssp126_tot_yearly, aes(x=year, y=sum, color="orange"), size = 0.5) + 
  geom_line(data = ipsl_ssp245_tot_yearly, aes(x=year, y=sum, color="red"), size = 0.5) +
  geom_line(data = ipsl_ssp370_tot_yearly, aes(x=year, y=sum, color = "blue"), size = 0.5) +
  geom_line(data = ipsl_ssp585_tot_yearly, aes(x=year, y=sum, color = "seagreen"), size = 0.5) +
  theme_minimal() + 
  #scale_y_continuous(breaks=seq(0,450,50)) +
  scale_x_continuous(breaks=seq(1850,2100,50)) + 
  # scale_x_date(date_breaks = "5 years", date_labels = "%Y") +
  ggtitle("Oahu, HI Yearly Total Precipitation, IPSL-CM6A-LR: 1850-2100") + 
  scale_color_manual(values = c("black", "orange", "red", "blue", "seagreen"), labels = c("historical", "ssp126", "ssp245", "ssp370", "ssp585")) +
  labs(x = "Year",
       y = "Precipitation (in)") +
  geom_smooth(data = ipsl_historical_tot_yearly, aes(x=year, y=sum), method='lm', formula = y~x, color = "black") +
  geom_smooth(data = ipsl_ssp126_tot_yearly, aes(x=year, y=sum), method='lm', formula = y~x, color = "orange") + 
  geom_smooth(data = ipsl_ssp245_tot_yearly, aes(x=year, y=sum), method='lm', formula = y~x, color = "red") + 
  geom_smooth(data = ipsl_ssp370_tot_yearly, aes(x=year, y=sum), method='lm', formula = y~x, color = "blue") + 
  geom_smooth(data = ipsl_ssp585_tot_yearly, aes(x=year, y=sum), method='lm', formula = y~x, color = "seagreen") +
  theme(legend.title=element_blank())+
    theme(legend.position="bottom")
# +
#   stat_regline_equation(label.y = 400) +
#   stat_cor(label.y = 400)

ipsl_yearly_sum_proj
```

## monthly totals 
```{r}
# all three ssps 
ipsl_monthly_sum_proj <- 
  ggplot() +
  geom_line(data = ipsl_historical_tot_monthly, aes(x=date, y=sum, color="black"), size = 0.5) +
  geom_line(data = ipsl_ssp126_tot_monthly, aes(x=date, y=sum, color="orange"), size = 0.5) + 
  geom_line(data = ipsl_ssp245_tot_monthly, aes(x=date, y=sum, color="red"), size = 0.5) +
  geom_line(data = ipsl_ssp370_tot_monthly, aes(x=date, y=sum, color = "blue"), size = 0.5) +
  geom_line(data = ipsl_ssp585_tot_monthly, aes(x=date, y=sum, color = "seagreen"), size = 0.5) +
  theme_minimal() + 
  scale_y_continuous(breaks=seq(0,300,50)) +
  #scale_x_continuous(breaks=seq(1850,2100,50)) + 
 scale_x_date(date_breaks = "50 years", date_labels = "%Y") +
  ggtitle("Oahu, HI Monthly Total Precipitation, IPSL-CM6A-LR: 1850-2100") + 
  scale_color_manual(values = c("black", "orange", "red", "blue", "seagreen"), labels = c("historical", "ssp126", "ssp245", "ssp370", "ssp585")) +
  labs(x = "Year",
       y = "Precipitation (in)") +
  # geom_smooth(data = bcc_historical_tot_monthly, aes(x=date, y=sum), method='lm', formula = y~x, color = "black") +
  # geom_smooth(data = bcc_ssp126_tot_monthly, aes(x=date, y=sum), method='lm', formula = y~x, color = "orange") +
  # geom_smooth(data = bcc_ssp245_tot_monthly, aes(x=date, y=sum), method='lm', formula = y~x, color = "red") +
  # geom_smooth(data = bcc_ssp370_tot_monthly, aes(x=date, y=sum), method='lm', formula = y~x, color = "blue") +
  # geom_smooth(data = bcc_ssp585_tot_monthly, aes(x=date, y=sum), method='lm', formula = y~x, color = "seagreen") +
  theme(legend.title=element_blank())+
    theme(legend.position="bottom")
  
  # stat_regline_equation(label.y = 200) +
  # stat_cor(label.y = 190)

ipsl_monthly_sum_proj
```
## separate historical total precip based on 30 yr time periods
```{r}
ipsl_historical_tot_1984_2014 <- ipsl_historical_tot %>% 
  filter(year %in% (1984:2014))
  
## separate projections based on 2020 to 2070
ipsl_ssp126_tot_2020_2050 <- ipsl_ssp126_tot %>% 
  filter(year %in% (2020:2050))

ipsl_ssp245_tot_2020_2050 <- ipsl_ssp245_tot %>% 
  filter(year %in% (2020:2050))

ipsl_ssp370_tot_2020_2050 <- ipsl_ssp370_tot %>% 
  filter(year %in% (2020:2050))

ipsl_ssp585_tot_2020_2050 <- ipsl_ssp585_tot %>% 
  filter(year %in% (2020:2050))


ipsl_daily_tot_period <- 
  ggplot() +
   geom_line(data = ipsl_historical_tot_1984_2014, aes(x=date, y=pcp, color="black"), size = 0.5) + 
  geom_line(data = ipsl_ssp126_tot_2020_2050, aes(x=date, y=pcp, color="orange"), size = 0.5) + 
  geom_line(data = ipsl_ssp245_tot_2020_2050, aes(x=date, y=pcp, color="red"), size = 0.5) +
  geom_line(data = ipsl_ssp370_tot_2020_2050, aes(x=date, y=pcp, color = "blue"), size = 0.5) +
  geom_line(data = ipsl_ssp585_tot_2020_2050, aes(x=date, y=pcp, color = "seagreen"), size = 0.5) + 
  theme_minimal() + 
  scale_y_continuous(breaks=seq(0,100,10), limits = c(0, 100)) +
  scale_x_date(date_breaks = "3 years", date_labels = "%Y") +
  ggtitle("Oahu, HI Daily Total Precipitation, IPSL-CM6A-LR: 1984-2050") +
   labs(x = "Year",
       y = "Precipitation (in)") +
  scale_color_manual(values = c("black", "orange", "red", "blue", "seagreen"), labels = c("historical", "spp126", "ssp245", "ssp370", "ssp585")) +
   theme(legend.title=element_blank())+
    theme(legend.position="bottom") +
    theme(text = element_text(size=14))

ipsl_daily_tot_period

# ggsave('ipsl_daily_tot_period.png', ipsl_daily_tot_period, width = 12, height = 8, units = "in")
```

## historical 50 year period graph
```{r}
ipsl_daily_tot_period_hist <- 
  ggplot() +
  geom_line(data = ipsl_historical_tot_1984_2014, aes(x=date, y=pcp, color="black"), size = 0.5) +  
  theme_minimal() + 
  scale_y_continuous(breaks=seq(0,50,10)) +
  scale_x_date(date_breaks = "3.5 years", date_labels = "%Y") +
  ggtitle("Oahu, HI Daily Total Precipitation, IPSL-CM6A-LR, historical: 1984-2014 ") +
   labs(x = "Year",
       y = "Precipitation (in)") +
  scale_color_manual(values = c("black"), labels = c("historical")) + 
 theme(legend.title=element_blank()) +
    theme(legend.position="bottom")

ipsl_daily_tot_period_hist
```


## ratio of daily total precipitation in the future vs. historical periods
```{r}
## merge the data frames together 
ipsl_daily_tot_prec_merge1 <- ipsl_ssp126_tot_2020_2050 %>% right_join(ipsl_ssp245_tot_2020_2050, by = c("date")) 
ipsl_daily_tot_prec_merge2 <- ipsl_daily_tot_prec_merge1 %>%  right_join(ipsl_ssp370_tot_2020_2050, by = c("date"))
ipsl_daily_tot_prec_merge_fin <- ipsl_daily_tot_prec_merge2 %>% right_join(ipsl_ssp585_tot_2020_2050,by = c("date")) %>% 
  select(date, year.x, month.x, day.x, pcp.x, pcp.y, pcp.x.x, pcp.y.y) %>% 
  rename(year = year.x,
         month = month.x,
         day = day.x,
         ssp126_pcp=pcp.x,
         ssp245_pcp=pcp.y,
         ssp370_pcp = pcp.x.x ,
         ssp585_pcp = pcp.y.y )

ipsl_historical_tot_1984_2014_rename <- ipsl_historical_tot_1984_2014 %>% 
  rename(hist_date = date,
         hist_pcp = pcp,
         hist_year = year,
         hist_month = month,
         hist_day = day)

ipsl_merge_his_proj <- cbind(ipsl_daily_tot_prec_merge_fin, ipsl_historical_tot_1984_2014_rename) 

ipsl_ratio_df <- ipsl_merge_his_proj %>% 
  mutate(ssp126_hist = ssp126_pcp/hist_pcp,
         ssp245_hist = ssp245_pcp/hist_pcp,
         ssp370_hist = ssp370_pcp/hist_pcp,
         ssp585_hist = ssp585_pcp/hist_pcp) %>% 
  select(date, ssp126_hist,ssp245_hist, ssp370_hist, ssp585_hist) %>% 
  pivot_longer(cols = ssp126_hist:ssp585_hist,
               names_to = c("ssp"),
               values_to = "fut_hist_ratio")

ggplot(data = ipsl_ratio_df, aes(x=date, y=fut_hist_ratio,  group=ssp)) +
  geom_boxplot() +
  facet_wrap(~ssp)
  #scale_y_continuous(limits = c(0,0.0e+00))
```


## look at analogeous storm event we will use for swmm 2009-10-11
```{r}
ssp_names <- c(
                    `ssp126_hist` = "ssp 1-2.6",
                    `ssp245_hist` = "ssp 2-4.5",
                    `ssp370_hist` = "ssp 3-7.0",
                    `ssp585_hist` = "ssp 5-8.5"
                    )

ipsl_merge_his_proj_filt <- ipsl_merge_his_proj %>% 
  filter(month == "10") %>% 
  filter(day == "11") %>% 
  mutate(ssp126_hist = ssp126_pcp/hist_pcp,
         ssp245_hist = ssp245_pcp/hist_pcp,
         ssp370_hist = ssp370_pcp/hist_pcp,
         ssp585_hist = ssp585_pcp/hist_pcp) %>% 
  select(date, year, month, day, ssp126_hist,ssp245_hist, ssp370_hist, ssp585_hist) %>% 
  pivot_longer(cols = ssp126_hist:ssp585_hist,
               names_to = c("ssp"),
               values_to = "fut_hist_ratio")


proj_hist_oct11_IPSL <- ggplot(data = ipsl_merge_his_proj_filt, aes(x=year, y=fut_hist_ratio,  group=year)) +
  geom_boxplot() +
  facet_wrap(~ssp, labeller = as_labeller(ssp_names)) +
  theme_minimal() +
  labs(x="Year", 
       y="Future Projection/Historical Daily Total Ratio (Oct 11)", 
       caption = "Data Source: IPSL-CM6A-LR") +
  scale_y_continuous(limits = c(0,30)) +
    theme(text = element_text(size=15))

proj_hist_oct11_IPSL

# ggsave('proj_hist_oct11_IPSL.png', proj_hist_oct11_IPSL, width = 12, height = 10, units = "in")
                     
```

## December 19, 2010 = Wailupe storm that we will use to do CC analysis with. Scrapping Kuliouou Oct 11th, 2009 storm, given we're having issues getting accurate SWMM output that correlated to the stream gage data
```{r}
ipsl_merge_his_proj_filt_dec19 <- ipsl_merge_his_proj %>%
  filter(month == "12") %>%
  filter(day == "19") %>%
  group_by(month,day) %>%
  mutate(ssp126_pcp_avg = mean(ssp126_pcp),
         ssp245_pcp_avg = mean(ssp245_pcp),
         ssp370_pcp_avg = mean(ssp370_pcp),
         ssp585_pcp_avg = mean(ssp585_pcp),
         hist_pcp_avg = mean(hist_pcp)) %>%
  select(month,day,ssp126_pcp_avg,ssp245_pcp_avg,ssp370_pcp_avg, ssp585_pcp_avg, hist_pcp_avg ) %>%
  distinct()


ipsl_merge_his_proj_filt_dec19 <- ipsl_merge_his_proj_filt_dec19 %>%
  mutate(ssp126_hist = ssp126_pcp_avg/hist_pcp_avg,
         ssp245_hist = ssp245_pcp_avg/hist_pcp_avg,
         ssp370_hist = ssp370_pcp_avg/hist_pcp_avg,
         ssp585_hist = ssp585_pcp_avg/hist_pcp_avg) %>%
  select(month, day, ssp126_hist,ssp245_hist, ssp370_hist, ssp585_hist) %>%
  pivot_longer(cols = ssp126_hist:ssp585_hist,
               names_to = c("ssp"),
               values_to = "fut_hist_ratio")

proj_hist_dec19_ipsl <- ggplot(data = ipsl_merge_his_proj_filt_dec19, aes(x=ssp, y=fut_hist_ratio)) +
  geom_point(size=2) +
  theme_minimal() +
  labs(x="Shared Socioeconomic Pathways (ssp)",
       y="Future Projection/Historical Daily Total Average Ratio (Dec 19)",
       caption = "Climate Model: IPSL-CM6A-LR") +
    theme(text = element_text(size=15)) +
  scale_x_discrete(labels=c("ssp126_hist" = "ssp 1-2.6",
                            "ssp245_hist" = "ssp 2-4.5",
                              "ssp370_hist" = "ssp 3-7.0",
                            "ssp585_hist" = "ssp 5-8.5")) + 
  scale_y_continuous(limits = c(0,3.5))

proj_hist_dec19_ipsl

# ggsave('proj_hist_dec19_ipsl.png', proj_hist_dec19_ipsl, width = 12, height = 10, units = "in")

```


## merging all climate model future/historical ratios (note: you will need to have run climate_change_analysis.Rmd for all climate models) to plot box plot
```{r}
futhis_ratio_combine <- merge_his_proj_filt_dec19 %>%
  rbind(bcc_merge_his_proj_filt_dec19, ipsl_merge_his_proj_filt_dec19) %>% 
  mutate(model = c("CNRM"))

futhis_ratio_combine$model[c(5:8)] = "BCC"
futhis_ratio_combine$model[c(9:12)] = "IPSL"

combine_ratio_plot <- ggplot(data = futhis_ratio_combine, aes(x=ssp, y=fut_hist_ratio)) +
  geom_boxplot() +
  geom_point(aes(color=model))+
  theme_minimal() +
  labs(x="Shared Socioeconomic Pathways (ssp)",
       y="Future Projection/Historical Daily Total Average Ratio (Dec 19)",
       caption = "CMIP6 Models: BCC-CSM2-MR, CNRM-ESM2-1, IPSL-CM6A-LR") +
    theme(text = element_text(size=15)) +
  scale_x_discrete(labels=c("ssp126_hist" = "ssp 1-2.6",
                            "ssp245_hist" = "ssp 2-4.5",
                              "ssp370_hist" = "ssp 3-7.0",
                            "ssp585_hist" = "ssp 5-8.5")) + 
  scale_y_continuous(limits = c(0,3.5)) +
  scale_color_discrete(name = "Model") +
    theme(legend.position="bottom")

combine_ratio_plot

#ggsave('combined_ratio_plot.png', combine_ratio_plot, width = 12, height = 10, units = "in")

```

